---
title: "Lession 4_Explore One Variable"
output: html_document
---

```{r}
library(ggplot2) 
data(diamonds)
```

## R Markdown

```{r}
summary(diamonds)
```

## Create a histogram of the price of
all the diamonds in the diamond data set.

```{r}
qplot(x = price, data = diamonds)
d1 <- NA
diamonds$d1 <- ifelse(diamonds$price < 500, 1, 0)
diamonds$d1 <- factor(diamonds$d1)
summary(diamonds$d1)

d2 <- NA
diamonds$d2 <- ifelse(diamonds$price < 250, 1, 0)
diamonds$d2 <- factor(diamonds$d2)
summary(diamonds$d2)

d3 <- NA
diamonds$d3 <- ifelse(diamonds$price == 15000, 1, 0)
diamonds$d3 <- factor(diamonds$d3)
summary(diamonds$d3)
```

## Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.
```{r}
qplot(x = price, data = diamonds, xlim = c(0, 15000))
qplot(x = price, data = diamonds) +
  scale_x_continuous(limits = c(0, 15000), breaks = seq(0, 15000, 1000))
ggsave('priceHistogram.png')


price1 <- qplot(x = price, data = diamonds) +
  scale_x_continuous(limits = c(0, 3000), breaks = seq(0, 3000, 100))

price2 <- qplot(x = price, data = diamonds) +
  scale_x_continuous(limits = c(3000, 6000), breaks = seq(3000, 6000, 100))

price3 <- qplot(x = price, data = diamonds) +
  scale_x_continuous(limits = c(6000, 9000), breaks = seq(6000, 9000, 100))

price4 <- qplot(x = price, data = diamonds) +
  scale_x_continuous(limits = c(9000, 12000), breaks = seq(9000, 12000, 100))

price5 <- qplot(x = price, data = diamonds) +
  scale_x_continuous(limits = c(12000, 15000), breaks = seq(12000, 15000, 100))

library(gridExtra)

# arrange plots in grid
grid.arrange(price1, price2, price3, price4, price5, ncol = 3)


qplot(x = price, data = diamonds) +
  scale_x_continuous(limits = c(0, 15000), breaks = 0:1000) +
  facet_wrap(~cut, ncol = 3)

by(diamonds$price, diamonds$cut, summary)
by(diamonds$price, diamonds$cut, summary, digits = max(getOption('digits')))
```
## Run the code below in R Studio to generate
# the histogram as a reminder.
```{r}
qplot(x = price, data = diamonds) + facet_grid(~cut, scales="free", space="free")

qplot(x = price, data = diamonds) + facet_grid(~cut, scales="free_y")

```
## Create a histogram of price per carat
# and facet it by cut.
```{r}
qplot(x = price / carat, data = diamonds, binwidth = 50) +
  scale_x_continuous(limits = c(0,15000), breaks = seq(0, 15000, 3000)) +
  facet_wrap(~cut, ncol = 3)
```
## Box Plots
need to add coord_cartesian()
For boxplots, we use the "y" parameter for the continuous data and the "x" parameter for the categorical data.
```{r}
qplot(x = cut, y = price,
      data = subset(diamonds, !is.na(cut)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 15000))

qplot(x = color, y = price,
      data = subset(diamonds, !is.na(color)),
      geom = 'boxplot') 

ggsave('priceHistogram.png')
```
#IQR
```{r}
by(diamonds$price,diamonds$color, summary)
IQR(subset(diamonds, color=='D')$price)
IQR(subset(diamonds, color=='J')$price)

```
## Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.
```{r}
qplot(x = price / carat, data = diamonds, binwidth = 50) +
  scale_x_continuous(limits = c(1000,9000), breaks = seq(1000, 9000, 1000)) +
  facet_wrap(~color, ncol = 2)

```
# Frequency Polygon
```{r}
qplot(x = carat, data = subset(diamonds, !is.na(carat)), 
      geom = 'freqpoly', binwidth = 0.05)
library(tidyr)
library(dplyr)
```
# Gapminder website
```{r}
library(ggplot2)
qplot(x = 2003, data = indicator_IT_TEL_TOTL_P2_xls)
sum(indicator_IT_TEL_TOTL_P2_xls$`1975.0`)

Aus <- subset(indicator_IT_TEL_TOTL_P2_xls, X__1 == 'Australia')
Chn <- subset(indicator_IT_TEL_TOTL_P2_xls, X__1 == 'China')
```
# Facebook Friends' birthdays
```{r}
getwd()
list.files()
fb <- read.csv('Udacity/EDA/birthdaysExample.csv', sep ='\t')
names(fb)

fb$dates <- as.Date(fb$dates, format = "%m/%d/%y")
fb$dates

month <- strftime(fb$dates, format = "%m")
month
day <- strftime(fb$dates, format = "%d")
day

qplot(month, xlab = 'Months of the month', ylab = 'Count', main = 'Month Frequency')
ggsave('MonthFrequency.png')

qplot(day, xlab = 'Days of the month', ylab = 'Count', main = 'Day Frequency')
ggsave('DayFrequency.png')
```